\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual All-Loop Integrands

### 2023-09-20 notes/2loopNLSM.tex
More comments from readthrough.

\immediate\write18{mkdir -p images}  %% Create `pgf-img` directory
\usetikzlibrary{external}  
\immediate\write18{mkdir -p images}  %% Create `pgf-img` directory
\tikzexternalize[                     %% Activate externalization
 prefix=images/,                    %% Avoid cluttering the directory
 system call={                       %% Use lualatex in system call
    pdflatex \tikzexternalcheckshellescape -halt-on-error -shell-escape -interaction=batchmode -jobname="\image" "\texsource" || rm "\image.pdf"
 },
]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

TODO:\\
For now, let's put the things we need to check/compute here in the abstract. 
\begin{enumerate}
\item \ace{I have defined a \textcolor{ucp-color}{\texttt{ucp-color}}
    , \textcolor{cut-color}{\texttt{cut-color}},
    \textcolor{refl-color}{\texttt{refl-color}}, and
    \textcolor{jacobi-color}{\texttt{jacobi-color}} as aliases for
    other colors so that we don't have to keep changing colors in
    multiple locations when we want to change diagram color schemes.
    I think I have updated all of the diagrams to use the
    correctly-named color, but haven't changed any of the backing
    colors for the names.}
\end{enumerate}
}

@@

Faced with this obstruction, we study the multi-loop sector of both
NLSM and Yang-Mills in \cref{2loopBoot} using an ansatz based
color-dual bootstrap. In contrast to much of the available literature
on color-dual integrand construction at two-loop
\cite{Bern:2013yya,Mogull:2015adi,Johansson:2017bfl}, our results are
completely agnostic to the spacetime dimension; all the polarizations
and momenta appearing in our construction will remain formally
$D$-dimensional. This makes our methods and results particularly well
suited for algorithmically extracting rational terms from the
$D$-dependence of dimensionally regulated loop momenta. Prior to this work, one of the primary barriers in studying multi-loop color-dual integrands $D$-dimensionally is the explosion of terms that one encounters in generic ansatz-based constructions at high-loop order. We were able overcome this barrier by employing a custom solver, \texttt{FiniteFieldSolve},
which will soon be made public. \ace{This either needs an extra
  context sentence, or moved out of the intro.}

@@

\ace{End of Alex first readthrough}

\section{One-loop cubic construction}
\label{sec:cubic}
Here we will describe why constructing the full Yang-Mills kinematic algebra (outside self-dual sector) is generically hard, and why the same reason makes NLSM beyond one-loop hard as well. It has to do with how the vector state sum mixes with the four-point contact. Starting with the Yang-Mills Lagrangian,
\begin{equation}
\mathcal{L}^{\text{YM}} = -\frac{1}{2}(\partial_\mu A_\nu)^2 + f^{abc} \partial_\mu A^a_\nu A^b_\mu A^c_\nu + f^{abe}f^{ecd}A^a_\mu A^b_\nu  A^c_\mu A^d_\nu 
\end{equation}
we can express the YM three-point interaction in terms of the following vertex in Lorenz gauge
\begin{equation}
T^k_{ij} = (\varepsilon_i\varepsilon_j)(\varepsilon_k p_i).
\end{equation}
This kinematic generator is anti-symmetric in $1\leftrightarrow 2$ exchange, and from it we can construct the full Lorenz gauge Yang-Mills vertex by summing over cyclic permutations
\begin{equation}
V^{(123)}_{\text{YM}} = T^3_{12} +\text{cyc}(123).
\end{equation}
Contracting this with the vector state sum, we can see that the cubic Yang-Mills vertex does not satisfy the four-point Jacobi identity
\be\label{cubicJac}
{}_1\langle V^{2}_{\text{YM}}V^{3}_{\text{YM}}\rangle_4+\text{cyc}(234) = \varepsilon_{(12)}  \varepsilon_{(34)} (s_{13}-s_{23}) +\text{cyc}(234) \neq 0
\end{equation}
where $ \varepsilon_{(ij)} =  \varepsilon_{i}\cdot  \varepsilon_{j}$.
The remaining terms can be absorbed into the definition of the four-point kinematic numerator
\begin{equation}
{}_1\langle V^{2}_{\text{YM}+B}V^{3}_{\text{YM}+B}\rangle_4 = {}_1\langle V^{2}_{\text{YM}}V^{3}_{\text{YM}}\rangle_4 + s_{12} (\varepsilon_{(13)}\varepsilon_{(24)}-\varepsilon_{(14)}\varepsilon_{(23)})
\end{equation}
where $V_{\text{YM}+B}$ indicates the introduction of an additional two-form studied in \cite{Ben-Shahar:2022ixa} for constructing the NMHV Yang-Mills Lagrangian.
After the introduction of the 2-form, the numerators now satisfy the Jacobi identity
\be
{}_1\langle V^{2}_{\text{YM}+B}V^{3}_{\text{YM}+B}\rangle_4+\text{cyc}(234) =0
\ee
where the new cubic Lagrangian takes the form
\be
\mathcal{L}^{\text{YM}+B} = -\frac{1}{2}(\partial_\mu A_\nu)^2 - B_{\mu\nu}\Box\tilde{B}_{\mu\nu}+ f^{abc} \partial_\mu A^a_\nu A^b_\mu A^c_\nu+ f^{abc} (B_{\mu\nu}+\Box \tilde{B}_{\mu\nu})^a A^b_\mu A^c_\nu\, .
\ee 
Of course, higher multiplicity would require further redefinition of the three-point vertex to satisfy Jacobi identites on all internal edges. While introducing successively higher spin states could in principle work at tree level, we will see that it is not consistent with what we find at general loop order. We will comment on this in \sect{}. For now, we will study how to construct one-loop integrands consistent with the color-kinematics by isolating the cubic sector of the state sum above.

As can be seen above, the Jacobi identity of \cref{cubicJac} fails only in terms with two factors of polarization dot products, $(\varepsilon\varepsilon)^2$. Whereas, if we restricted ourselves to just considering factors with a single polarization dot product, the Jacobi identity of Yang-Mills would be satisfied \textit{off-shell} and in arbitrary spacetime dimensions
\begin{equation}
{}_1\langle V^{2}_{\text{YM}}V^{3}_{\text{YM}}\rangle_4\big|^{(\varepsilon\varepsilon)^1}+\text{cyc}(234) =0\,.
\end{equation}
Thus for our approach at one-loop, we aim to build color-dual integrands directly from kinematic factors of this form. The organizational principle underlying this construction can be understood in terms of a polarization decomposition of vector amplitudes introduced by one of the authors \cite{Pavao:2022kog}, 
\begin{equation}\label{eq:pureVecRABD}
A_{(\sigma)}^{\text{YM}} = \sum_{k=0}^{\lfloor |\sigma|/2\rfloor}\sum_{\rho \in S^{2|k}_{\sigma}}\varepsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
\end{equation}
\jm{Nic, could you give a quick recap of what all of the symbols mean in the equation above?}
By a simple mass-dimensioin argument, we know the tree-level cubic sector of Yang-Mills must be at $\mathcal{O}((\varepsilon\varepsilon)^1)$,
\begin{equation}
A_{(\text{tree})}^{\text{cubic}} = \sum_{\rho \in S^{2|1}_{\sigma}}\varepsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\, ,
\end{equation}
while at one-loop, the cubic sector corresponds to $\mathcal{O}((\varepsilon\varepsilon)^0)$ in polarization dot products,
\begin{equation}
A_{(\text{1-loop})}^{\text{cubic}} = \sum_{\rho \in S^{2|0}_{\sigma}}\varepsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
\end{equation}
When decomposed in this way, the polarization stripped building blocks must obey a set of Ward-identities between different ``helicity'' (or $(\varepsilon \varepsilon)$) sectors in order for the full amplitude to be gauge invariant,
 \begin{equation}\label{eq:GIrelA}
\Delta_{(\sigma)}^{(\rho)}\Big|_{\epsilon_i\rightarrow k_i} =-
\sum_{j \in \rho^c} (k_i \varepsilon_j)\Delta_{(\sigma )}^{(\rho\cup (ij))} \,.
\end{equation}
With this in hand, we can construct a Lagrangian description of the of the cubic sector for Yang-Mills, and will demonstrate that the resulting amplitudes with manifestly color-dual Feynman rules are equivalent to both SDYM and NLSM through one-loop. 
\subsection{Semi-abelian Yang-Mills theory}\label{semiYM}
As we demonstrate above, if we include factors of $(\varepsilon\varepsilon)^{n\geq 1}$ at tree-level, then we need to keep the Yang-Mills four-point contact for color-kinematics duality to be restored on-shell. However, the contrapositive is also true -- if we omit the four-point Yang-Mills vertex, then we only need terms that contribute to the manifestly color-dual cubic sector of the theory.

We can achieve this construction by splitting the gauge group of the full Yang-Mills action from $G$ to $G_L \times G_R$, where the full non-abelian gauge symmetry is restored only when we identify $G_L = G_R$. However, if we break the left gauge group down to $G_L \rightarrow U(1)^{N^2}$ and take the right gauge group to $G_R \rightarrow U(N)$, then we are left with a semi-abelian theory that is gauge invariant in the preferred $U(1)$-directions. We call this manifestly color-dual theory, \textit{semi-abelian Yang-Mills},
\begin{eBox}
\begin{equation}
\label{eq:semiYM}
\mathcal{L}^{\text{semi-YM}} = -\frac{1}{2}\text{tr}\left[\bar{F}_{\mu\nu}F^{\mu\nu}\right]
\end{equation}
\end{eBox}
where
\begin{equation}
\bar{F}^a_{\mu\nu} = \partial_\mu \bar{A}_\nu^a - \partial_\nu \bar{A}_\mu^a
\end{equation}
\begin{equation}
F^a_{\mu\nu} = \partial_\mu A_\nu^a - \partial_\nu A_\mu^a + f^{abc} A_\mu^b A_\nu^c .
\end{equation}
Here we can think of the abelian vector as a background field that sources on-shell $A_\mu$ currents. Indeed, as we describe in Appendix \cref{sec:CKLagrangians}, semi-abelian Yang-Mills theory is at the heart of YZ-theory \cite{Cheung:2016prv}, J-theory \cite{Cheung:2020djz,Cheung:2021zvb}, and self-dual Yang-Mills \cite{Monteiro2011pc}.
Specifically, semi-abelian YM is simply a clever reinterpretation of the Lagrangian obtained by integrating an auxiliary field into the J-theory equations of motion.
The Feynman rule for the cubic vertex is
\begin{equation}
\cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3) -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3).
\end{equation}
The propagator kinetic mixes between the two gauge fields and has the standard Feynman rule of $i/k^2$.
We can use the Ward identity of \cref{eq:GIrelA} to check that the amplitudes are indeed gauge invariant when the abelian polarizations are taken to be longitudinal, $\bar{\varepsilon} \rightarrow k$.

We can select out both of these manifestly cubic sectors by selecting only $\varepsilon^+\varepsilon^+$ in light-cone gauge (i.e. SDYM) and one-minus at tree-level OR what $J$-theory and $YZ$-theory select with their on-shell states. Indeed, the $D$-dimensional description of $J$-theory is just the following blocks in the expansion of Yang-Mills given in \cite{Pavao:2022kog}
\begin{equation}
A(...,\bar{A}_k,...)_{\text{tree}} = \sum_{j\neq k}\Delta^{(jk)}_{\text{tree}}
\end{equation}
and similarly so at one-loop
\begin{equation}
A(...)_{\text{1-loop}} \equiv \Delta^{(\varnothing)}_{\text{1-loop}}\, .
\end{equation}
To recover NLSM at one-loop, we need to extract the $D$-dependent part that corresponds to an internal $YY$-loop from extra-dimensional scalars
\begin{eBox}
\begin{equation}
A^{\text{NLSM}}_{\text{1-loop}} \equiv \partial_D \Delta^{(\varnothing)}_{\text{1-loop}}\big|_{\varepsilon\rightarrow k}\, .
\end{equation}
\end{eBox}
Why must we take the derivative with respect to $D$? After all, $J$-theory and $YZ$-theory both have well defined propagators for internal $\bar{J}J$ and $\bar{Z}Z$ propagators. However, as we noted in the previous section, all the $J$ and $Z$ states must be on-shell in order to produce NLSM amplitudes. Thus, the unitarity cuts of $J$ theory at one-loop will not produce NLSM amplitudes. However, the internal $YY$-loop is a valid forward limit, since as constructed YZ theory is valid for off-shell Y-particles, and thus they serve as a valid forward limit. We will construct these one-loop integrands below in the next section.

\subsection{One-loop color-dual integrands}\label{oneLoopCK}
Our first application of this theory for color-dual construction at loop level is for self-dual Yang-Mills (SDYM). At tree-level, the off-shell cubic vertices can be written in terms of light cone coordinates \cite{Monteiro2011pc}:
X(p,k) = p_u k_w-p_w k_u
At loop-level, this construction needs to be analytically continued to general dimension in order to apply dimensional regularization at one-loop. This can be done using the cubic semi-YM vertices of the previous section:
\cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= \mathcal{X}(k_1,k_2) = (\varepsilon_1\varepsilon_3)(\varepsilon_2 k_3) -  (\varepsilon_2\varepsilon_3)(\varepsilon_1 k_3)
Plugging in on-shell all-plus helicity states in light-cone gauge will yield precisely the 4D SDYM vertex, up to an unphysical phase. Of course, this has the advantage of permitting a $D$-dimensional construction of the one-loop integrand. Below we provide a four-point example:
\simplebox = g_{\text{YM}}^4 \langle \mathcal{X}(k_1,l_1)\mathcal{X}(k_2,l_2)\mathcal{X}(k_3,l_3)\mathcal{X}(k_4,l_4)\rangle
where $l_i = l+k_1+k_2+\cdots+ k_i$. In general, the $n$-gon vertex would be
\mathcal{N}^{\text{SDYM}}_{n\text{-gon}}= \nGonVector = g_{\text{YM}}^n \langle \mathcal{X}(k_1,l_1)\mathcal{X}(k_2,l_2)\cdots \mathcal{X}(k_n,l_n) \rangle
Due to the state-sum of internal loop factors, $\sum {\varepsilon_{(+l)}\varepsilon_{(-l)}}=D$, the integrand above depends explicitly on the spacetime dimension, $D$. By taking a derivative\footnote{Another way to understand this derivative is that it selects the large $D$ behavior of the integrand.}, we can recover the integrand needed for all-plus one-loop amplitudes with an internal scalar:
\frac{\partial}{\partial D} \mathcal{N}^{\text{SDYM}}_{n\text{-gon}}=\nGonScalar  = (2 g_{\text{YM}})^n (l_1 k_1)(l_2 k_2)\cdots (l_n k_n)
This integrand numerator with internal scalar loop is precisely what one would obtain from the Feynman rules of YZ-theory, absent the $XZ$ internal vector loop. For more background, we refer the reader to \cref{sec:CKLagrangians}. 
 \mathcal{N}^{\text{NLSM}}_{n\text{-gon}}=\left[\frac{\partial}{\partial D} \mathcal{N}^{\text{SDYM}}_{n\text{-gon}}\right]^{\epsilon \rightarrow k}=\nGonScalar =  f_\pi^{-n} [12][23]\cdots [n1]
Where we have define the antisymmetric kinematic variable, $[LR] = l_L^2 - l_R^2$
\subsection{Two-loop obstruction}\label{2LoopObstruction}
At two-loop, introducing terms that conspire with four-point contacts is unavoidable. At one-loop, we were able to avoid internal contractions of $\bar{A}_\mu A^\mu$ by selecting external states of appropriate weight. However, at two-loop when choosing all external $A_\mu$ states, the amplitude vanishes in semi-YM theory:
\mathcal{A}^{\text{semi-YM}}_{\text{2-loop}}(+,+,+,+)=0
In order to produce non-vanishing interactions, we would need to reintroduce $D$-dimensional vertices from the diagonal group of $G_L\times G_R$. Doing so will necessarily introduce internal contractions of $\bar{A}_\mu A^\mu$,
\mathcal{A}^{\text{YM}}_{\text{2-loop}}(+,+,+,+)=\doubleBoxObstructionA +\doubleBoxObstructionB+\cdots 
However, this immediately runs into the difficulty of introducing the four-point contact needed for color-kinematics to be satisfied on all internal edges. To do this prescriptively would require knowledge of the full kinematic algebra for Yang-Mills off-shell. Since this is presently unavailable, to approach the problem of two-loop construction, we instead employ an ansatz based approach. 
\section{Two-loop bootstrap}\label{2loopBoot}

@@

\subsection{Two-loop NLSM}
\label{sec:pions}

As we have seen, it is possible to coerce tree numerators into 1-loop numerators at any multiplicity for pions.
However, these methods cannot be reapplied to generate higher-loop numerators, which begs the question ``is it possible to find globally color-dual numerators beyond 1-loop''.
We will tackle this question for 4-point 2-loop scattering in NLSM before moving on to the same process in YM.

>>>>

\usetikzlibrary{external}  
\immediate\write18{mkdir -p images}  %% Create `pgf-img` directory
\tikzexternalize[                     %% Activate externalization
prefix=images/,                    %% Avoid cluttering the directory
system call={                       %% Use lualatex in system call
   pdflatex \tikzexternalcheckshellescape -halt-on-error -shell-escape -interaction=batchmode -jobname="\image" "\texsource" || rm "\image.pdf"
},
]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

TODO:\\
For now, let's put the things we need to check/compute here in the abstract. 
\begin{enumerate}
\item Settle on $n$ vs $N$ vs $\mathcal{N}$ as the symbol for
  numerator
\item \ace{I have defined a \textcolor{ucp-color}{\texttt{ucp-color}}
    , \textcolor{cut-color}{\texttt{cut-color}},
    \textcolor{refl-color}{\texttt{refl-color}}, and
    \textcolor{jacobi-color}{\texttt{jacobi-color}} as aliases for
    other colors so that we don't have to keep changing colors in
    multiple locations when we want to change diagram color schemes.
    I think I have updated all of the diagrams to use the
    correctly-named color, but haven't changed any of the backing
    colors for the names.}
\end{enumerate}
}

@@

Faced with this obstruction, we study the multi-loop sector of both
NLSM and Yang-Mills in \cref{2loopBoot} using an ansatz based
color-dual bootstrap. In contrast to much of the available literature
on color-dual integrand construction at two-loop
\cite{Bern:2013yya,Mogull:2015adi,Johansson:2017bfl}, our results are
completely agnostic to the spacetime dimension; all the polarizations
and momenta appearing in our construction will remain formally
$D$-dimensional. This makes our methods and results particularly well
suited for algorithmically extracting rational terms from the
$D$-dependence of dimensionally regulated loop momenta. Prior to this
work, one of the primary barriers in studying multi-loop color-dual
integrands $D$-dimensionally is the explosion of terms that one
encounters in generic ansatz-based constructions at high-loop
order. We were able overcome this barrier by employing a custom
solver, \texttt{FiniteFieldSolve}, which will soon be made
public.

@@

\section{One-loop cubic construction}
\label{sec:cubic}
Here we will describe why constructing the full Yang-Mills kinematic
algebra (outside the self-dual sector) is generically hard, and why
the same reason makes NLSM beyond one-loop hard as well. The major
obstacle stems from the vector state sum mixing with higher-point
contacts. Starting with the Yang-Mills Lagrangian,
\begin{equation}
  \mathcal{L}^{\text{YM}} = -\frac{1}{2}(\partial_\mu A_\nu)^2
  + f^{abc} \partial_\mu A^a_\nu A^b_\mu A^c_\nu
  + f^{abe}f^{ecd}A^a_\mu A^b_\nu  A^c_\mu A^d_\nu 
\end{equation}
we can express the YM three-point interaction in terms of the
following vertex in Lorenz gauge
\begin{equation}
T^k_{ij} = (\varepsilon_i\varepsilon_j)(\varepsilon_k p_i).
\end{equation}
This kinematic generator is anti-symmetric in $1\leftrightarrow 2$
exchange, and from it we can construct the full Lorenz gauge
Yang-Mills vertex by summing over cyclic permutations
\begin{equation}
V^{(123)}_{\text{YM}} = T^3_{12} +\text{cyc}(123).
\end{equation}
Contracting this with the vector state sum, we can see that the cubic
Yang-Mills vertex does not satisfy the four-point Jacobi identity
\begin{equation}
  \label{cubicJac}
  {}_1\langle V^{2}_{\text{YM}}V^{3}_{\text{YM}}\rangle_4+\text{cyc}(234) = \varepsilon_{(12)}  \varepsilon_{(34)} (s_{13}-s_{23}) +\text{cyc}(234) \neq 0
\end{equation}
where $ \varepsilon_{(ij)} = \varepsilon_{i}\cdot \varepsilon_{j}$.
The remaining terms can be absorbed into the definition of the
four-point kinematic numerator
\begin{equation}
{}_1\langle V^{2}_{\text{YM}+B}V^{3}_{\text{YM}+B}\rangle_4 = {}_1\langle V^{2}_{\text{YM}}V^{3}_{\text{YM}}\rangle_4 + s_{12} (\varepsilon_{(13)}\varepsilon_{(24)}-\varepsilon_{(14)}\varepsilon_{(23)})
\end{equation}
where $V_{\text{YM}+B}$ indicates the introduction of an additional
two-form studied in Ref.~\cite{Ben-Shahar:2022ixa} for constructing the
NMHV Yang-Mills Lagrangian.  After the introduction of the 2-form, the
numerators now satisfy the Jacobi identity
\begin{equation}
  {}_1\langle
  V^{2}_{\text{YM}+B}V^{3}_{\text{YM}+B}\rangle_4+\text{cyc}(234) =0
\end{equation}
where the new cubic Lagrangian takes the form
\begin{equation}
\mathcal{L}^{\text{YM}+B} = -\frac{1}{2}(\partial_\mu A_\nu)^2 -
B_{\mu\nu}\Box\tilde{B}_{\mu\nu}+ f^{abc} \partial_\mu A^a_\nu A^b_\mu
A^c_\nu+ f^{abc} (B_{\mu\nu}+\Box \tilde{B}_{\mu\nu})^a A^b_\mu
A^c_\nu\, .
\end{equation}
Of course, higher multiplicity would require further redefinition of
the three-point vertex to satisfy Jacobi identites on all internal
edges. While introducing successively higher spin states could in
principle work at tree level, we will see that it is not consistent
with what we find at general loop order. We will comment on this in
\cref{sec:Discussion} \ace{Is this the sec you are thinking of?}. For
now, we will study how to construct one-loop integrands consistent
with the color-kinematics by isolating the cubic sector of the state
sum above.

As can be seen above, the Jacobi identity of \cref{cubicJac} fails
only in terms with two factors of polarization dot products,
$(\varepsilon\varepsilon)^2$. Whereas if we restricted ourselves to
just considering factors with a single polarization dot product, the
Jacobi identity of Yang-Mills would be satisfied \textit{off-shell}
and in arbitrary spacetime dimensions
\begin{equation}
{}_1\langle V^{2}_{\text{YM}}V^{3}_{\text{YM}}\rangle_4\big|^{(\varepsilon\varepsilon)^1}+\text{cyc}(234) =0\,.
\end{equation}
\ace{Should we mention that this is closely related to MHV vs NMHV?}
Thus for our approach at one-loop, we aim to build color-dual
integrands directly from kinematic factors of this form. The
organizational principle underlying this construction can be
understood in terms of a polarization decomposition of vector
amplitudes introduced by one of the authors \cite{Pavao:2022kog},
\begin{equation}\label{eq:pureVecRABD}
A_{(\sigma)}^{\text{YM}} = \sum_{k=0}^{\lfloor |\sigma|/2\rfloor}\sum_{\rho \in S^{2|k}_{\sigma}}\varepsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
\end{equation}
\jm{Nic, could you give a quick recap of what all of the symbols mean
  in the equation above?}  By a simple mass-dimensioin argument, we
know the tree-level cubic sector of Yang-Mills must be at
$\mathcal{O}((\varepsilon\varepsilon)^1)$,
\begin{equation}
A_{(\text{tree})}^{\text{cubic}} = \sum_{\rho \in S^{2|1}_{\sigma}}\varepsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\, ,
\end{equation}
while at one-loop, the cubic sector corresponds to
$\mathcal{O}((\varepsilon\varepsilon)^0)$ in polarization dot
products,
\begin{equation}
A_{(\text{1-loop})}^{\text{cubic}} = \sum_{\rho \in S^{2|0}_{\sigma}}\varepsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
\end{equation}
When decomposed in this way, the polarization stripped building blocks
must obey a set of Ward-identities between different ``helicity'' (or
$(\varepsilon \varepsilon)$) sectors in order for the full amplitude
to be gauge invariant,
 \begin{equation}\label{eq:GIrelA}
\Delta_{(\sigma)}^{(\rho)}\Big|_{\epsilon_i\rightarrow k_i} =-
\sum_{j \in \rho^c} (k_i \varepsilon_j)\Delta_{(\sigma )}^{(\rho\cup (ij))} \,.
\end{equation}
With this in hand, we can construct a Lagrangian description of the of
the cubic sector for Yang-Mills, and will demonstrate that the
resulting amplitudes with manifestly color-dual Feynman rules are
equivalent to both SDYM and NLSM through one-loop.

\subsection{Semi-abelian Yang-Mills theory}\label{semiYM}
As we demonstrate above, if we include factors of
$(\varepsilon\varepsilon)^{n\geq 1}$ at tree-level, then we need to
keep the Yang-Mills four-point contact for color-kinematics duality to
be restored on-shell. However, the contrapositive is also true -- if
we omit the four-point Yang-Mills vertex, then we only need terms that
contribute to the manifestly color-dual cubic sector of the theory.

We can achieve this construction by splitting the gauge group of the
full Yang-Mills action from $G$ to $G_L \times G_R$, where the full
non-abelian gauge symmetry is restored only when we identify
$G_L = G_R$. However, if we break the left gauge group down to
$G_L \rightarrow U(1)^{N^2}$ and take the right gauge group to
$G_R \rightarrow U(N)$, then we are left with a semi-abelian theory
that is gauge invariant in the preferred $U(1)$-directions. We call
this manifestly color-dual theory \textit{semi-abelian Yang-Mills},
\begin{eBox}
\begin{equation}
\label{eq:semiYM}
\mathcal{L}^{\text{semi-YM}} = -\frac{1}{2}\text{tr}\left[\bar{F}_{\mu\nu}F^{\mu\nu}\right]
\end{equation}
\end{eBox}
where
\begin{equation}
\bar{F}^a_{\mu\nu} = \partial_\mu \bar{A}_\nu^a - \partial_\nu \bar{A}_\mu^a
\end{equation}
\begin{equation}
F^a_{\mu\nu} = \partial_\mu A_\nu^a - \partial_\nu A_\mu^a + f^{abc} A_\mu^b A_\nu^c .
\end{equation}
Here we can think of the abelian vector as a background field that
sources on-shell $A_\mu$ currents. Indeed, as we describe in Appendix
\cref{sec:CKLagrangians}, semi-abelian Yang-Mills theory is at the
heart of YZ-theory \cite{Cheung:2016prv}, J-theory
\cite{Cheung:2020djz,Cheung:2021zvb}, and self-dual Yang-Mills
\cite{Monteiro2011pc}.  Specifically, semi-Abelian YM is simply a
clever reinterpretation of the Lagrangian obtained by integrating an
auxiliary field into the J-theory equations of motion.  The Feynman
rule for the cubic vertex is
\begin{equation}
  \cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3)
  -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3).
\end{equation}
The kinetic term mixes the two gauge fields and has the standard
Feynman rule of $i/k^2$.  We can use the Ward identity of
\cref{eq:GIrelA} to check that the amplitudes are indeed gauge
invariant when the abelian polarizations are taken to be longitudinal,
$\bar{\varepsilon} \rightarrow k$.

We can select out both \ace{which both?} of these manifestly cubic
sectors by selecting only $\varepsilon^+\varepsilon^+$ in light-cone
gauge (i.e. SDYM) and one-minus at tree-level OR what $J$-theory and
$YZ$-theory select with their on-shell states. Indeed, the
$D$-dimensional description of $J$-theory is just the following blocks
in the expansion of Yang-Mills given in Ref.~\cite{Pavao:2022kog}
\begin{equation}
A(...,\bar{A}_k,...)_{\text{tree}} = \sum_{j\neq k}\Delta^{(jk)}_{\text{tree}}
\end{equation}
and similarly so at one-loop
\begin{equation}
A(...)_{\text{1-loop}} \equiv \Delta^{(\varnothing)}_{\text{1-loop}}\, .
\end{equation}
To recover NLSM at one-loop, we need to extract the $D$-dependent part
that corresponds to an internal $YY$-loop from extra-dimensional
scalars
\begin{eBox}
\begin{equation}
A^{\text{NLSM}}_{\text{1-loop}} \equiv \partial_D \Delta^{(\varnothing)}_{\text{1-loop}}\big|_{\varepsilon\rightarrow k}\, .
\end{equation}
\end{eBox}
Why must we take the derivative with respect to $D$? After all,
$J$-theory and $YZ$-theory both have well defined propagators for
internal $\bar{J}J$ and $\bar{Z}Z$ propagators. However, as we discuss
in \cref{sec:CKLagrangians} \ace{I think this is where we do this
  now}, all the $J$ and $Z$ states must be on-shell in order to
produce NLSM amplitudes. Thus, the unitarity cuts of $J$ theory at
one-loop will not produce NLSM amplitudes. However, the internal
$YY$-loop is a valid forward limit for producing NLSM amplitudes,
since as constructed YZ theory mathces to NLSM for off-shell
Y-particles. We will construct these one-loop integrands in the
next section.

\subsection{One-loop color-dual integrands}\label{oneLoopCK}
Our first application of this theory for color-dual construction at
loop level is for self-dual Yang-Mills (SDYM). At tree-level, the
off-shell cubic vertices can be written in terms of light cone
coordinates \cite{Monteiro2011pc}:
\begin{equation}
X(p,k) = p_u k_w-p_w k_u
\end{equation}
At loop-level, this construction needs to be analytically continued to
general dimension in order to apply dimensional regularization at
one-loop. This can be done using the cubic semi-YM vertices of the
previous section:
\begin{equation}
\cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= \mathcal{X}(k_1,k_2) = (\varepsilon_1\varepsilon_3)(\varepsilon_2 k_3) -  (\varepsilon_2\varepsilon_3)(\varepsilon_1 k_3)
\end{equation}
Plugging in on-shell all-plus helicity states in light-cone gauge will
yield precisely the 4D SDYM vertex, up to an unphysical phase \ace{Can
  we provide an example?}. Of course, this has the advantage of
permitting a $D$-dimensional construction of the one-loop
integrand. Below we provide a four-point example:
\begin{equation}
\simplebox = g_{\text{YM}}^4 \langle \mathcal{X}(k_1,l_1)\mathcal{X}(k_2,l_2)\mathcal{X}(k_3,l_3)\mathcal{X}(k_4,l_4)\rangle
\end{equation}
where $l_i = l+k_1+k_2+\cdots+ k_i$. In general, the $n$-gon vertex would be
\begin{equation}
\mathcal{N}^{\text{SDYM}}_{n\text{-gon}}= \nGonVector = g_{\text{YM}}^n \langle \mathcal{X}(k_1,l_1)\mathcal{X}(k_2,l_2)\cdots \mathcal{X}(k_n,l_n) \rangle
\end{equation}
Due to the state-sum of internal loop factors,
$\sum {\varepsilon_{(+l)}\varepsilon_{(-l)}}=D$, the integrand above
depends explicitly on the spacetime dimension, $D$. By taking a
derivative\footnote{Another way to understand this derivative is that
  it selects the large $D$ behavior of the integrand.}, we can recover
the integrand needed for all-plus one-loop amplitudes with an internal
scalar:
\begin{equation}
\frac{\partial}{\partial D} \mathcal{N}^{\text{SDYM}}_{n\text{-gon}}=\nGonScalar  = (2 g_{\text{YM}})^n (l_1 k_1)(l_2 k_2)\cdots (l_n k_n)
\end{equation}
This integrand numerator with internal scalar loop is precisely what
one would obtain from the Feynman rules of YZ-theory, absent the $XZ$
internal vector loop. For more background, we refer the reader to
\cref{sec:CKLagrangians}.
\begin{equation}
 \mathcal{N}^{\text{NLSM}}_{n\text{-gon}}=\left[\frac{\partial}{\partial D} \mathcal{N}^{\text{SDYM}}_{n\text{-gon}}\right]^{\epsilon \rightarrow k}=\nGonScalar =  f_\pi^{-n} [12][23]\cdots [n1]
\end{equation}
Where we have define the antisymmetric kinematic variable, $[LR] = l_L^2 - l_R^2$

\subsection{Two-loop obstruction}\label{2LoopObstruction}
At two-loop, introducing terms that conspire with four-point contacts
is unavoidable. At one-loop, we were able to avoid internal
contractions of $\bar{A}_\mu A^\mu$ by selecting appropriate external
states. However, at two-loop when choosing all external $A_\mu$
states, the amplitude vanishes in semi-YM theory:
\begin{equation}
\mathcal{A}^{\text{semi-YM}}_{\text{2-loop}}(+,+,+,+)=0
\end{equation}
In order to produce non-vanishing interactions, we would need to
reintroduce $D$-dimensional vertices from the diagonal group of
$G_L\times G_R$. Doing so will necessarily introduce internal
contractions of $\bar{A}_\mu A^\mu$,
\begin{equation}
\mathcal{A}^{\text{YM}}_{\text{2-loop}}(+,+,+,+)=\doubleBoxObstructionA +\doubleBoxObstructionB+\cdots 
\end{equation}
However, this immediately runs into the difficulty of introducing the
four-point contact needed for color-kinematics to be satisfied on all
internal edges. To do this prescriptively would require knowledge of
the full kinematic algebra for Yang-Mills off-shell. Since this is
presently unavailable, to approach the problem of two-loop
construction, we instead employ an ansatz based approach.

\section{Two-loop bootstrap}
\label{2loopBoot}
\ace{Lets bulk this section intro up a bit}

@@

\subsection{Two-loop NLSM}
\label{sec:pions}

As we have seen, it is possible to coerce tree numerators into 1-loop
numerators at any multiplicity for pions.  However, these methods
cannot be reapplied to generate higher-loop numerators, which begs the
question ``is it possible to find globally color-dual numerators
beyond 1-loop?''  We will tackle this question for 4-point 2-loop
scattering in NLSM before moving on to the same process in YM.
